---
title: "Household_cluster_size"
author: "Dan Weinberger"
date: "3/3/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Some issues
1) We don't know the date of infection; only the date when a person is tested
2) Vaccine could prevent symptoms but not infection, so unvaccinated HH member might get tested first (incorrectly appearing as index)
3) As time goes on and more people are vaccinated, te index cases are more likely to be younger and might not be comparable to index cases from earlier in study
4) Decision to get tested could be influenced by vaccination
5) Decision to get tested is influenced by presence of an infected HH member
6) If people were infected/positive earlier in epidemic, need to exclude them as susceptibles

Longini paper: https://academic.oup.com/aje/article/115/5/736/153401?login=true
"A great deal of information concerning community transmission is contained in the comparison between households that had no infections and those experiencing at least one infection."

*B* as the probability that a susceptible household member escapes being infected from the community during the period of observation

*Q* as the probability a household member escapes infectious contact from a single infected household member during his entire infectious period

the probability that **j** of **k** initial susceptibles become infected in a particular household during the period of observation can be found.

m_jk gives probability that j of k susceptibles become infected

```{r}
ds <- matrix(c(340,197,84,60,25,11,3,2,1,0,
                 164,104,60,29,15,6,4,2,0,0,
                 0,57,57,25,9,4,0,3,0,1,
                 0,0,27,11,10,3,3,3,0,0,
                 0,0,0,7,1,0,2,0,0,0,
                 0,0,0,0,1,1,3,0,0,0,
                 rep(0,5),1,rep(0,4),
                 rep(0,6),1,1,0,0,
                 rep(0,30)),
               nrow=11,byrow=T)

max.hh.size= ncol(ds) # Maximum household size = # of columns 
choose_kj_mat <- matrix(0, nrow=max.hh.size, ncol=max.hh.size)
for(k in 2: max.hh.size){
  for(j in 1:k){
    choose_kj_mat[k,j] <- choose(k,j)
  }
}

```


```{r}
source('./R/longini_jags.R')
```

```{r}
##############################################################
#Model Fitting
##############################################################
inits1=list(".RNG.seed"=c(123), ".RNG.name"='base::Wichmann-Hill')
inits2=list(".RNG.seed"=c(456), ".RNG.name"='base::Wichmann-Hill')
inits3=list(".RNG.seed"=c(789), ".RNG.name"='base::Wichmann-Hill')



##############################################
#Model Organization
##############################################
model_spec<-textConnection(longini_jags)
model_jags<-jags.model(model_spec, 
                       inits=list(inits1,inits2, inits3),
                       data=list( 
                                 'ds'=ds,
                                 'max.hh.size'=max.hh.size,
                                 'choose_kj_mat'=choose_kj_mat
                                 ),
                       n.adapt=10000, 
                       n.chains=3)

```

